---
title: "Data Cleaning"
format: pdf
editor: visual
---

### Data from TMDB API and IMDB non-commercial


```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(gender)
library(readxl)
library(data.table)
library(mice)
film <- read.csv("/Users/yvette/Desktop/film/tmdb/TMDB_movie_dataset_v11.csv")
imdb_rating <- read.csv("/Users/yvette/Desktop/film/imdb data noncomercial/title_ratings_transformed.csv")
```

### merge the fiml data with imdb_ratings

```{r}
# merge the fiml data with imdb_ratings
film <- film[order(film$id), ]
head(film)

cleaned_film <- film %>%
  select(-backdrop_path, -homepage, -overview, -poster_path, -tagline, -keywords)
cleaned_film[cleaned_film == ""] <- NA

na_count_per_column <- cleaned_film %>%
  summarise_all(~ sum(is.na(.)))

cleaned_film <- cleaned_film %>%
  filter(!is.na(title) & !is.na(release_date) & 
           !is.na(imdb_id) & !is.na(production_companies)& 
           !(revenue == 0 & budget == 0) & !(imdb_id == "None"))

na_count_per_column <- cleaned_film %>%
  summarise_all(~ sum(is.na(.)))

duplicates_in_film <- cleaned_film %>%
  count(id) %>%
  filter(n > 1)

print(duplicates_in_film)

cleaned_film <- cleaned_film %>%
  group_by(id) %>%
  slice_head(n = 1) %>%
  ungroup()

merged_df <- cleaned_film %>%
  left_join(imdb_rating, by = c("imdb_id" = "tconst"))

film_with_ratings <- merged_df %>%
  rename(imdb_rating = averageRating,
         imdb_votes_count = numVotes)

id_list <- as.list(cleaned_film$id)
id_df <- data.frame(id = unlist(id_list), stringsAsFactors = FALSE)
write.csv(id_df, file = "/Users/yvette/Desktop/film/id_df.csv", row.names = FALSE)
```

```{r}
#| echo: false
#| warning: false
film <- film_with_ratings %>%
  filter(status == 'Released') %>%
  mutate(imdb_votes_count = ifelse(is.na(imdb_votes_count), 0, imdb_votes_count),
         adult = as.logical(adult),
         adult = ifelse(adult == TRUE, 1, 0),
         english = ifelse(original_language == "en", 1, 0)) %>%
  mutate(genre = str_extract(genres, "^[^,\\s]+")) %>%
  dplyr::select(id,title,release_date,runtime,adult,genre,english,revenue,budget,imdb_votes_count,imdb_id)
```


```{r}
#| echo: false
#| warning: false
#| include: false
oscar <- read_excel("/Users/yvette/Desktop/film/List_of_Academy_Award–nominated_films.xlsx")
number_of_procom<- read.csv("/Users/yvette/Desktop/film/separated_production_companies.csv")
runtime <- read_tsv('/Users/yvette/Desktop/film/imdb data noncomercial/title.basics.tsv') 
origin_country <- read.csv("/Users/yvette/Desktop/film/df.csv")
```

### Genres, Runtime, Artistic, Number of production companies, title word count,origin country,holiday dummy

```{r}
#| echo: false
#| warning: false
#| include: false
# genre
genre_frequency <- film %>%
  count(genre, sort = TRUE)

film <- film %>%
  mutate(genres_category = case_when(
    genre == "Drama" ~ "Drama",
    genre == "Comedy" ~ "Comedy",
    genre == "Horror" ~ "Horror",
    genre == "Action" ~ "Action",
    genre == "Documentary" ~ "Documentary",
    TRUE ~ "Other"  
  ))

runtime <- as_tibble(runtime)
merged_data <- runtime %>%
  dplyr::select(tconst, runtimeMinutes) %>%      
  rename(imdb_id = tconst) %>%           
  inner_join(select(film, imdb_id, runtime), 
             by = "imdb_id")  

# runtime 
merged_data$runtimeMinutes[merged_data$runtimeMinutes == "\\N"] <- NA
na_count <- sum(is.na(merged_data$runtimeMinutes))
count_zeros <- sum(merged_data$runtime == 0, na.rm = TRUE)
count_different <- sum(merged_data$runtimeMinutes != merged_data$runtime, na.rm = TRUE)

merged_data <- merged_data %>%
  mutate(runtime = as.numeric(runtime),
         runtimeMinutes = as.numeric(runtimeMinutes)) %>%
  mutate(runtime = if_else(runtime == 0, runtimeMinutes, runtime))

# title
film <- film %>%
  mutate(title_word_count = str_count(title, "\\S+"))

# Oscar
oscar <- oscar %>%
  mutate(artistic = 1)
film <- film %>%
  left_join(oscar %>% select(Film, artistic), by = c("title" = "Film")) %>%
  mutate(artistic = ifelse(is.na(artistic), 0, artistic))
count_award <- sum(film$artistic == 1, na.rm = TRUE)

# Number of production companies
number_of_procom[number_of_procom == ""] <- NA
number_of_procom$number_of_com <- rowSums(!is.na(number_of_procom[, -1]))
number_of_procom <- number_of_procom %>%
  select(id, number_of_com)

film <- film %>%
  left_join(number_of_procom, by = "id") 

# origin_country
origin_country$origin_country <- gsub("\\[|\\]|'", "", origin_country$origin_country)
origin_country <- origin_country %>%
  rename(id = movie_id) %>%
  distinct()
film <- inner_join(film, origin_country, by = "id")
film <- film %>%
  mutate(origin_country_us = ifelse(str_detect(str_split_fixed(origin_country, ",", 2)[, 1], "US"), 1,0))

# Holidays
film_holiday <- film %>%
  select(id, release_date) %>%
  mutate(release_date = date(release_date)) %>%
  mutate(
    week = isoweek(release_date),                   
    year = year(release_date)                      
  ) %>%
  mutate(
    labor_day_week = isoweek(as.Date(paste0(year, "-09-01")) + days(1 - wday(as.Date(paste0(year, "-09-01"))))), # First Monday in September
    memorial_day_week = isoweek(as.Date(paste0(year, "-05-31")) - days(wday(as.Date(paste0(year, "-05-31"))) - 1)), # Last Monday in May
    thanksgiving_week = isoweek(as.Date(paste0(year, "-11-01"))) + 3, # 4th Thursday in November
    new_year_week = isoweek(as.Date(paste0(year, "-01-01"))),       # New Year's Day
    christmas_week = isoweek(as.Date(paste0(year, "-12-25"))),     # Christmas Day
    valentine_week = isoweek(as.Date(paste0(year, "-02-14"))),     # Valentine's Day
    independence_week = isoweek(as.Date(paste0(year, "-07-04"))),  # Independence Day
    halloween_week = isoweek(as.Date(paste0(year, "-10-31")))      # Halloween week
  ) %>%
  mutate(
    new_year_dummy = if_else(week == new_year_week, 1, 0),
    valentine_dummy = if_else(week == valentine_week, 1, 0),
    independence_dummy = if_else(week == independence_week, 1, 0),
    labor_day_dummy = if_else(week == labor_day_week, 1, 0),
    memorial_day_dummy = if_else(week == memorial_day_week, 1, 0),
    thanksgiving_dummy = if_else(week == thanksgiving_week, 1, 0),
    christmas_dummy = if_else(week == christmas_week, 1, 0),
    halloween_dummy = if_else(week == halloween_week, 1, 0)
  ) %>%
  mutate(
    holiday_dummy = if_else(
      new_year_dummy == 1 | valentine_dummy == 1 | independence_dummy == 1 | 
      labor_day_dummy == 1 | memorial_day_dummy == 1 | 
      thanksgiving_dummy == 1 | christmas_dummy == 1 | 
      halloween_dummy == 1, 1, 0
    )
  ) %>%
  select(id,new_year_dummy,valentine_dummy,independence_dummy,labor_day_dummy,memorial_day_dummy,thanksgiving_dummy,christmas_dummy,halloween_dummy,holiday_dummy)

holiday_movies <- sum(film_holiday$holiday_dummy) #4541

film$holiday_dummy <- film_holiday$holiday_dummy
```




```{r}
#| echo: false
#| warning: false
#| include: false
cast <- read.csv("/Users/yvette/Desktop/film/cast_data.csv")
crew <- read.csv("/Users/yvette/Desktop/film/crew_data.csv")
```

### Gender Imputation (gender_ratio_female_acting, gender_ratio_female_crew, gender_ratio_female_tota, number_of_crew, number_of_departments) and Cast and Crew merged

```{r}
#| echo: false
#| warning: false
#| include: false
num_unique_values <- length(unique(cast$movie_id))

cast <- cast %>%
  mutate(gender = factor(gender))
sorted_cast <- cast %>%
  arrange(movie_id, order) %>%
  select(-cast_id,-character) %>%
  mutate(department = "Acting", job = "Actor/Actress")
sorted_cast[sorted_cast== ""] <- NA

sorted_crew <- crew %>%
  mutate(gender = factor(gender)) %>%
  arrange(movie_id, department) %>%
  select(-credit_id) %>%
  group_by(movie_id, department) %>%
  mutate(order = row_number() - 1) %>%
  ungroup()
sorted_crew$department[sorted_crew$department == "Actors"] <- "Crew"
cast_and_crew_full <- bind_rows(sorted_cast, sorted_crew)

gender_zero <- cast_and_crew_full[cast_and_crew_full$gender == 0, ] %>%
  select(id,name) %>%
  separate(name, into = c("first_name", "last_name"), 
           sep = " ", extra = "merge", fill = "right") %>%
  mutate(full_name = paste(first_name, last_name))

gender_results <- gender(gender_zero$first_name)
gender_results_unique <- gender_results[!duplicated(gender_results$name), ]
gender_zero_r <- gender_zero %>%
  left_join(gender_results_unique %>% select(name, gender), by = c("first_name" = "name"))
gender_zero_r <- gender_zero_r %>%
  mutate(gender = case_when(
    gender == "male" ~ 2,
    gender == "female" ~ 1,
    gender == "other" ~ 3,
    is.na(gender) ~ 0,  # Handle NA values
    TRUE ~ NA_real_  # Handle any unexpected values
  )) %>%
  rename(imput_gender = gender)
gender_imput<- cast_and_crew_full %>%
  left_join(gender_zero_r %>% select(id, imput_gender), by = "id")%>%
  mutate(imput_gender = ifelse(is.na(imput_gender), 0, imput_gender), 
    gender = as.numeric(as.character(gender)),  
    imput_gender = as.numeric(as.character(imput_gender)),  
    gender = gender + imput_gender) %>%
  select(-imput_gender)

summary_table <- gender_imput %>%
  group_by(movie_id) %>%
  summarise(number_of_crew = n_distinct(name),
    number_of_departments = n_distinct(department)) %>%
  select(movie_id, number_of_crew, number_of_departments)

# Aggregating Acting and Other Departments
gender_ratios_summary <- gender_imput %>%
  mutate(department_group = ifelse(department == "Acting", "Acting", "Other_Departments")) %>%
  group_by(movie_id, department_group) %>%
  summarize(
    num_male = sum(gender == 2, na.rm = TRUE),
    num_female = sum(gender == 1, na.rm = TRUE),
    num_other = sum(gender == 3, na.rm = TRUE),
    num_unknown = sum(gender == 0, na.rm = TRUE),
    gender_ratio_female = num_female /(num_female + num_male + num_other + num_unknown),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = department_group,
    values_from = c(num_male, num_female, num_other, gender_ratio_female),
    names_glue = "{.value}_{department_group}"
  ) %>%
  rename(
    num_male_acting = num_male_Acting,
    num_female_acting = num_female_Acting,
    num_male_crew = num_male_Other_Departments,
    num_female_crew = num_female_Other_Departments,
    num_other_acting = num_other_Acting,
    num_other_crew = num_other_Other_Departments,
    gender_ratio_female_acting = gender_ratio_female_Acting,
    gender_ratio_female_crew = gender_ratio_female_Other_Departments
  ) %>%
  mutate(gender_ratio_female_total = (num_female_acting + num_female_crew) /
      (num_female_acting + num_female_crew + num_male_acting + num_male_crew + num_other_acting + num_other_crew),
      across(everything(), ~ replace_na(., 0))) %>%
  select(movie_id, gender_ratio_female_acting, gender_ratio_female_crew, gender_ratio_female_total)

filtered_cast_and_crew <- gender_imput %>%
  filter(movie_id %in% film_final$movie_id)
write_csv(filtered_cast_and_crew,"cast_and_crew_full.csv")

film <- film %>%
  left_join(summary_table, by = c("id" = "movie_id")) %>%
  left_join(gender_ratios_summary, by = c("id" = "movie_id"))
```

### Imputation and Dependent variables

```{r}
#| echo: false
#| warning: false
#| include: false
# imputation budget
film_imputation<- film %>%
  mutate(
    revenue = ifelse(revenue == 0, NA, revenue),
    budget = ifelse(budget == 0, NA, budget)
  )
value_imputed_budget <- data.frame(
  original = film_imputation$budget,
  imputed_mean = replace(film_imputation$budget, is.na(film_imputation$budget), mean(film_imputation$budget, na.rm = TRUE)),
  imputed_median = replace(film_imputation$budget, is.na(film_imputation$budget), median(film_imputation$budget, na.rm = TRUE))
)

h1 <- ggplot(value_imputed_budget, aes(x = original)) +
  geom_histogram(fill = "#ad1538", color = "#000000", position = "identity") +
  ggtitle("Original distribution") +
  theme_classic()
h2 <- ggplot(value_imputed_budget, aes(x = imputed_mean)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("Mean-imputed distribution") +
  theme_classic()
h3 <- ggplot(value_imputed_budget, aes(x = imputed_median)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("Median-imputed distribution") +
  theme_classic()

film_imputation_numeric <- film_imputation %>%
  select(budget,number_of_crew)
md.pattern(film_imputation_numeric)
mice_imputed_budget <- data.frame(
  original = film_imputation_numeric$budget,
  imputed_pmm = complete(mice(film_imputation_numeric, method = "pmm"))$budget,
  imputed_cart = complete(mice(film_imputation_numeric, method = "cart"))$budget,
  imputed_lasso = complete(mice(film_imputation_numeric, method = "lasso.norm"))$budget
)


h4 <- ggplot(mice_imputed_budget, aes(x = imputed_pmm)) +
  geom_histogram(fill = "#15ad4f", color = "#000000", position = "identity") +
  ggtitle("pmm-imputed distribution") +
  theme_classic()
h5 <- ggplot(mice_imputed_budget, aes(x = imputed_cart)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("cart-imputed distribution") +
  theme_classic()
h6 <- ggplot(mice_imputed_budget, aes(x = imputed_lasso)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("lasso-imputed distribution") +
  theme_classic()



film_imputation <- film_imputation %>%
  mutate(imputed_pmm_budget = mice_imputed_budget$imputed_pmm)

# imputation revenue
value_imputed <- data.frame(
  original = film_imputation$revenue,
  imputed_mean = replace(film_imputation$revenue, is.na(film_imputation$revenue), mean(film_imputation$revenue, na.rm = TRUE)),
  imputed_median = replace(film_imputation$revenue, is.na(film_imputation$revenue), median(film_imputation$revenue, na.rm = TRUE))
)

h2_1 <- ggplot(value_imputed, aes(x = original)) +
  geom_histogram(fill = "#ad1538", color = "#000000", position = "identity") +
  ggtitle("Original distribution") +
  scale_x_sqrt() +
  theme_classic()
h2_2 <- ggplot(value_imputed, aes(x = imputed_mean)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("Mean-imputed distribution") +
  scale_x_sqrt() +
  theme_classic()
h2_3 <- ggplot(value_imputed, aes(x = imputed_median)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("Median-imputed distribution") +
  scale_x_sqrt() +
  theme_classic()

film_imputation_numeric <- film_imputation %>%
  select(revenue,imdb_votes_count)
md.pattern(film_imputation_numeric)
mice_imputed <- data.frame(
  original = film_imputation_numeric$revenue,
  imputed_pmm = complete(mice(film_imputation_numeric, method = "pmm"))$revenue,
  imputed_cart = complete(mice(film_imputation_numeric, method = "cart"))$revenue,
  imputed_lasso = complete(mice(film_imputation_numeric, method = "lasso.norm"))$revenue
)

h2_4 <- ggplot(mice_imputed, aes(x = imputed_pmm)) +
  geom_histogram(fill = "#15ad4f", color = "#000000", position = "identity") +
  ggtitle("pmm-imputed distribution") +
  scale_x_sqrt() +
  theme_classic()
h2_5 <- ggplot(mice_imputed, aes(x = imputed_cart)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("cart-imputed distribution") +
  scale_x_sqrt() +
  theme_classic()
h2_6 <- ggplot(mice_imputed, aes(x = imputed_lasso)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("lasso-imputed distribution") +
  scale_x_sqrt() +
  theme_classic()



film_imputation <- film_imputation %>%
  mutate(imputed_cart_revenue = mice_imputed$imputed_cart,
         box = ifelse(imputed_cart_revenue > imputed_pmm_budget, 1, 0),
         box_artistic = ifelse(box == 1 & artistic == 1, 1, 0))

film_final <- film_imputation %>%
  select(id,title_word_count,runtime,adult,english,origin_country_us,holiday_dummy,genres_category,release_date,number_of_crew,number_of_departments,number_of_com,gender_ratio_female_total, gender_ratio_female_acting, gender_ratio_female_crew, imputed_pmm_budget,imputed_cart_revenue,box,artistic,box_artistic) %>%
  distinct(id, .keep_all = TRUE) %>%
  filter(complete.cases(.))

# Continuous profit
film_final <- film_final %>%
  mutate(profit = imputed_cart_revenue - imputed_pmm_budget,
    log_profit = log(pmax(film_final$profit, 0) + 1),
    log_revenue = log(pmax(film_final$imputed_cart_revenue, 0) + 1),
    log_budget = log(pmax(film_final$imputed_pmm_budget, 0) + 1)) 
film_final <- film_final %>%
  mutate(
    adult = as.factor(adult),
    english = as.factor(english),
    genres_category = as.factor(genres_category),
    origin_country_us = as.factor(origin_country_us),
    holiday_dummy = as.factor(holiday_dummy)
  )  %>%
  rename(movie_id = id) 

write.csv(film_final, "film_final.csv", row.names = FALSE)
```

